Climate indicators

### 2023-02-27 DRAFT_climate_indicators.Rmd
Loop through dir

@@

I will remove the _band_ dimension.
```{r}
x <- x %>% filter(band==1) %>% adrop()
```

@@

I'll add some extra seasonal variation between to the mix
```{r}
x$Attribute_A[,,2] <- x$Attribute_A[,,1]*1.2
x$Attribute_A[,,3] <- x$Attribute_A[,,1]*1.4
x$Attribute_A[,,4] <- x$Attribute_A[,,1]*1.1
```

@@

    -   Join all data into one data cube (´stars:c´)

1.  Calculate reference values

    -   Aggregate across reference years to get median and sd values `st_apply`

    -   Join with existing data cube

2.  Calculate indicator values

    -   Normalize climate variable at the individual grid cell level using the three reference values `mutate %>% across`

@@

    -   Aggregate across 5 year time steps to smooth out random interannual variation and leav climate signal

@@

### Collate variable data series

@@

Then for each iteration I need to get the year as well
```{r}
(year_temp <- year(st_get_dimension_values(x, "time")[1]))
```

Then I can create a time interval object
```{r}
start <- ym(paste(year_temp, start_month, sep="-"))
end <-  ym(paste(year_temp, end_month, sep="-"))
(myInterval <- interval(start, end))
```

@@

```{r, fig.cap="There is no performance increas by making before aggregating"}
autoplot(
  microbenchmark(
No_masking =  x_aug %>%
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names)),
Masking =  x_masked %>%
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names))
 ), times=30
)

@@

```{r}
temp_names <- year(st_get_dimension_values(x_aug, "time")[1])
x_summerMean <-   x_aug %>%
  filter(band ==1) %>% 
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names))
```

@@

#### Aggregate across years

For this step we need some more dummy data from different years.

@@

#### Export
I should export the reference data at this point (write to file) as we don't need to recalculate this every time we want to update the indicator values.

@@

Calculate diference between indicator and reference
We don't want to include this in ts, but as a temp file.
```{r}
tsX <- ts1
```

>>>>

## --------------------------------------------------

@@

I will remove the *band* dimension.

```{r}
x <- x %>% filter(band==1) %>% adrop()
```

@@

I'll add some extra seasonal variation between to the mix

```{r}
x$Attribute_A[,,2] <- x$Attribute_A[,,1]*1.2
x$Attribute_A[,,3] <- x$Attribute_A[,,1]*1.4
x$Attribute_A[,,4] <- x$Attribute_A[,,1]*1.1
```

@@

    -   Join all data into one data cube (`stars:c`)

2.  Calculate reference values

    -   Aggregate across reference years to get median and sd values (`st_apply)`

    -   Join with existing data cube (`stars:c`)

3.  Calculate indicator values

    -   Normalize climate variable at the individual grid cell level using the three reference values (`mutate(across()))`

@@

    -   Aggregate across 5 year time steps to smooth out random inter-annual variation and leave climate signal

@@

### 1. Collate variable data series

@@

Then for each iteration I need to get the year as well

```{r}
(year_temp <- year(st_get_dimension_values(x, "time")[1]))
```

Then I can create a time interval object

```{r}
start <- ym(paste(year_temp, start_month, sep="-"))
end <-  ym(paste(year_temp, end_month, sep="-"))
(myInterval <- interval(start, end))
```

@@

```{r withMasking, fig.cap="Compareing computation time before and after masking of raster data. There is no performance increase by masking before aggregating."}
autoplot(
  microbenchmark(
No_masking =  x_aug %>%
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names)),
Masking =  x_masked %>%
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names))
 ), times=30
)

@@

Since there is no increase in performance from masking, as we see in Fig. \@ref(fig:withMasking) then it basically means we it is slower to mask beforehand, since the masking was not past of the benchmark assessment.

So, this is my solution:

```{r}
x_summerMean <-   x_aug %>%
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names))
```

The *v\_* follows the naming convention in Norway.

@@

Here's the whole first step of the conceptual workflow in action, on the dummy data.

```{r}

# Setting up the parameters
dir <- substr(getwd(), 1,2)
path <- ifelse(dir == "C:",
       "P:/", 
       "/data/P-Prosjekter2/")
path2 <- paste0(path, 
                "41201785_okologisk_tilstand_2022_2023/data/climate_indicators/dummy_data")
start_month <-  "06"
end_month <- "08"
myFiles <- list.files(path2, full.names = T)
summerTemp_fullSeries <- NULL

# Looping though the files in the directory
system.time({
for(i in 1:length(myFiles)){

  temp <- readRDS(myFiles[i]) %>%
    select(Attribute_A)

  year_temp <- year(st_get_dimension_values(temp, "time")[1])
  start <- ym(paste(year_temp, start_month, sep="-"))
  end <-  ym(paste(year_temp, end_month, sep="-"))
  myInterval <- interval(start, end)

  temp <- temp %>%
    filter(time %within% myInterval) %>%
    st_apply(1:2, mean) %>%
    setNames(paste0("v_",year_temp))

  summerTemp_fullSeries <- c(temp, summerTemp_fullSeries)
  rm(temp) # same computation time with and without this function, but more tidy this way
}

# Turn the attributes into a dimension and rename the new attribute
summerTemp_fullSeries <- summerTemp_fullSeries %>%
  merge(name="Year") %>%
  setNames("Attribute_A")
})
```

And this is the result.

```{r}
ggplot()+
  geom_stars(data = summerTemp_fullSeries) +
  coord_equal() +
  theme_void() +
  scale_fill_viridis_c(option = "D") +  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0))+
  facet_wrap(~Year)
```

### 2. Calculate reference values
#### Aggregate across reference years

For this step we need some more dummy data from different years. I made data sets y1 to y5 above.

@@

#### Export

I should export the reference data at this point (write to file) as we don't need to recalculate this every time we want to update the indicator values.

@@

Calculate diference between indicator and reference We don't want to include this in ts, but as a temp file.

```{r}
tsX <- ts1
```
